---
title: "Project 1-Topic 3 Report"
author: "Hongjie Liu, Jiajun Tao, Shaohan Chen"
date: "02/27/2023"
header-includes:
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{subfigure}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gridExtra)
library(tidyverse)
library(patchwork)
library(dplyr)
```


## Background

When dealing with high-dimensional data

## Statistical methods to be studied

### Stepwise Forward Selection

Step-wise forward selection 

### LASSO

LASSO regression 

## Objective

Several parameters 


## Simulation

Data is generated

## Experiment Settings and Scenario



## Model Evaluation

### Evaluation Metrics
In this project, we define the true predictors as positive and null predictors as negative.\par
For signal identification, we use the following five metrics to compare the two models:\par
- Complexity: The number of selected predictors in the model
- Sensitivity: $\frac{TP}{TP+FN}$
- Specificity: $\frac{TN}{TN+FP}$
- F1-score: $\frac{2\cdot sensitivity\cdot precision}{sensitivity + precision}$
- Accuracy: $\frac{TP+TN}{TP+TN+FP+FN}$\par

For parameter estimation, we use the following to metrics to compare the two models:\par
- RMSE: $\sqrt{\frac{1}{p}\sum_{i=1}^p (\hat{\beta_i}-\beta_i)^2}$
- Variance: $\sqrt{\frac{1}{p}\sum_{i=1}^p (\hat{\beta_i}-\bar{\beta_i})^2}$

### Signal Indentification Performance



Complexity of the models is indicated by the number of selected predictors. We can see that in high dimensional scenario(when n=100), Forward selection model tends to select lots of predictors and Lasso tends to select very few. One interesting thing is that, if we increase the ratio of strong predictors(i.e. more strong predictors), Lasso also tends to select more predictors too. When it comes to normal scenario, Forward selection still tends to select more predictors than Lasso, but the discrepancy is smaller than high dimensional case, and will be further narrowed down with n or the ratio of strong predictors increasing. And as n increases, the number of selected parameters of both models are closer to the true number 40.

As for overall classification performance, if in high dimensional scenario, Forward selection tends to be very assertive and much better at identifying weak signals, leading to an extremely high sensitivity but low specificity. Lasso in turn tends to be very conservative and much better at identifying null signals, leading to an extremely high specificity but low sensitivity. Like high dimensional case, Lasso becomes more sensitive and not that assertive when ratio of strong predictors increases. Based on the above plot, we can conclude that both models does not perform too well based on F1-score and accuracy, because they are very radical and tend to identify most of the predictors either as positive or negative, but far away from the truth.\par
For normal scenario, both models become less radical under normal scenarios, but Forward selection is still more sensitive and assertive than Lasso, while Lasso has higher specificity and more conservative. Both models perform better on those metrics with n increasing. Overall, Lasso and Forward selection has similar F1-score and accuracy performance. But when there are more strong predictors, Lasso performs obviously better than Forward selection.

About the classification performance of different signals. Under all n values, both models perfectly identify the strong signals. In high dimensional scenario, Forward selection performs much better on identifying weak predictors while Lasso performs much better on identifying null predictors, that’ s why we see the high sensitivity of forward selection and high specificity of Lasso in previous section. When there are more strong predictors, Lasso also performs better on selecting weak predictors. When it comes to normal scenario, Forward selection is still better at selecting null and Lasso is better at selecting weak predictors. But the discrepancy is smaller compared with high-dimensional data, and will continue be smaller as n increases. When there are more strong predictors, Lasso performs much better at selecting weak-but-correlated signals.

### Parameter Estimatio Performance

In high dimensional scenario, Lasso performs much better than forward selection, with obvious much lower and centered RMSE and also lower variance. When it comes to normal scenario. Though when n=500, Lasso outperforms forward selection, but with n increasing, forward selection starts to outperform Lasso model on RMSE and variance. And overall, Lasso tends to perform better when there are more strong signals. If there are more strong predictors, the variance is also larger


### Effect of Missing Weak Predictors

On the other hand

## Discussions

*Limitation*

*Future Work* 

## Reference


\pagebreak

## Appendix

```{r, include=FALSE, warning = FALSE}
# Load parameter data
param_1 = read_csv("parameter_estimate.csv")
param_2 = read_csv("parameter_estimate2.csv")

param_1 =
  param_1 %>%
  filter(method %in% c("forward", "lasso_lmfit")) %>%
  mutate(method = 
           case_when(
             method == "forward" ~ "forward",
             method == "lasso_lmfit" ~ "lasso"
           ))
param_2 =
  param_2 %>%
  filter(method %in% c("forward", "lasso_lmfit")) %>%
  mutate(method = 
           case_when(
             method == "forward" ~ "forward",
             method == "lasso_lmfit" ~ "lasso"
           ))

param_estimate = rbind(param_1, param_2)


# Output identification table
fit_identify_result = 
  param_estimate %>% 
    mutate(
      n = factor(n),
      ratio = ifelse(ratio == 145, "1:4:5", "3:3:4"),
      notnull = ifelse(type == "null", 0, 1),
      notnull_est = ifelse(abs(estimate) > 0, 1, 0),
      identify =
        case_when(
          notnull == 1 & notnull_est == 1 ~ "TP",
          notnull == 1 & notnull_est == 0 ~ "FN",
          notnull == 0 & notnull_est == 1 ~ "FP",
          notnull == 0 & notnull_est == 0 ~ "TN"
        )
    ) %>% 
    group_by(sim_time, n, p, ratio, method) %>% 
    summarize(
      number_of_parameters = sum(identify == "TP") + sum(identify == "FP"),
      sensitivity = sum(identify == "TP")/(sum(identify == "TP") + sum(identify == "FN")),
      specificity = sum(identify == "TN")/(sum(identify == "TN") + sum(identify == "FP")),
      precision = sum(identify == "TP")/(sum(identify == "TP") + sum(identify == "FP")),
      accuracy = (sum(identify == "TP") + sum(identify == "TN")) /
        (sum(identify == "TP") + sum(identify == "FP") + sum(identify == "TN") + sum(identify == "FN")),
      F1_score = 2 * precision * sensitivity / (precision + sensitivity)
    )


# Output identification table by signal
fit_identify_result_type = 
  param_estimate %>% 
    mutate(
      n = factor(n),
      ratio = ifelse(ratio == 145, "1:4:5", "3:3:4"),
      notnull = ifelse(type == "null", 0, 1),
      notnull_est = ifelse(abs(estimate) > 0, 1, 0),
      identify =
        case_when(
          notnull == 1 & notnull_est == 1 ~ "TP",
          notnull == 1 & notnull_est == 0 ~ "FN",
          notnull == 0 & notnull_est == 1 ~ "FP",
          notnull == 0 & notnull_est == 0 ~ "TN"
        )
    ) %>% 
    group_by(sim_time, n, p, ratio, method, type) %>% 
    summarize(
      number_of_parameters = sum(identify == "TP") + sum(identify == "FP"),
      sensitivity = sum(identify == "TP")/(sum(identify == "TP") + sum(identify == "FN")),
      specificity = sum(identify == "TN")/(sum(identify == "TN") + sum(identify == "FP")),
      precision = sum(identify == "TP")/(sum(identify == "TP") + sum(identify == "FP")),
      accuracy = (sum(identify == "TP") + sum(identify == "TN")) /
        (sum(identify == "TP") + sum(identify == "FP") + sum(identify == "TN") + sum(identify == "FN")),
      F1_score = 2 * precision * sensitivity / (precision + sensitivity)
    )


p = 100
ratio = c(1, 4, 5)
ps = floor(0.4*p*ratio[1]/sum(ratio))
pwbc = floor(0.4*p*ratio[2]/sum(ratio))
pwai = floor(0.4*p*ratio[3]/sum(ratio))
beta = c(rep(20, ps), rep(0.5, pwbc + pwai), rep(0, p - ps - pwbc - pwai))
# output the beta estimation table of ratio 1
fit_beta_result_ratio1 = 
  param_1 %>% 
    mutate(
      n = factor(n),
      ratio = ifelse(ratio == 145, "1:4:5", "3:3:4"),
      num = case_when(
  type == "strong" ~ num,
  type == "wbc" ~ num + ps,
  type == "wai" ~ num + ps + pwbc,
  type == "null" ~ num + ps + pwbc + pwai),
      se = (estimate - beta[num])^2,
      var = (estimate - mean(estimate))^2
    ) %>% 
    group_by(sim_time, n, p, ratio, method) %>% 
    summarize(
      RMSE = sqrt(mean(se)),
      Variance = mean(var)
    )

# output the beta estimation table of ratio2
p = 100
ratio = c(3, 3, 4)
ps = floor(0.4*p*ratio[1]/sum(ratio))
pwbc = floor(0.4*p*ratio[2]/sum(ratio))
pwai = floor(0.4*p*ratio[3]/sum(ratio))
beta = c(rep(20, ps), rep(0.5, pwbc + pwai), rep(0, p - ps - pwbc - pwai))

fit_beta_result_ratio2 = 
  param_2 %>% 
    mutate(
      n = factor(n),
      ratio = ifelse(ratio == 145, "1:4:5", "3:3:4"),
      num = case_when(
  type == "strong" ~ num,
  type == "wbc" ~ num + ps,
  type == "wai" ~ num + ps + pwbc,
  type == "null" ~ num + ps + pwbc + pwai),
      se = (estimate - beta[num])^2,
      var = (estimate - mean(estimate))^2
    ) %>% 
    group_by(sim_time, n, p, ratio, method) %>% 
    summarize(
      RMSE = sqrt(mean(se)),
      Variance = mean(var)
    )

fit_beta_result_ratio = 
  rbind(fit_beta_result_ratio1, fit_beta_result_ratio2)
```

```{r, include  = FALSE}
# compare model complexity

# compare sensitivity
g_1 = ggplot(fit_identify_result, aes(x = method, y = number_of_parameters, color = n)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Model Complexity",x = "method", y = "Number of Selected Parameters") +
  facet_grid(. ~ ratio)
```


```{r, include  = FALSE}
# compare all signals on four metrics

# compare sensitivity
g1 = ggplot(fit_identify_result, aes(x = method, y = sensitivity, color = n)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Sensitivity",x = "method", y = "sensitivity") +
  facet_grid(. ~ ratio)

# compare specificity
g2 = ggplot(fit_identify_result, aes(x = method, y = specificity, color = n)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Specificity",x = "method", y = "specificity") +
  facet_grid(. ~ ratio)

# compare F1-score
g3 = ggplot(fit_identify_result, aes(x = method, y = F1_score, color = n)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "F1_score",x = "method", y = "F1_score") +
  facet_grid(. ~ ratio)

# compare accuracy
g4 = ggplot(fit_identify_result, aes(x = method, y = accuracy, color = n)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Accuracy",x = "method", y = "accuracy") +
  facet_grid(. ~ ratio)


# compare different signals

# First ratio
g5 = ggplot(fit_identify_result_type, aes(x = method, y = sensitivity, color = type)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Sensitivity",x = "method", y = "sensitivity") +
  facet_grid(ratio ~ n)

g6 = ggplot(fit_identify_result_type, aes(x = method, y = specificity, color = type)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Specificity",x = "method", y = "specificity") +
  facet_grid(ratio ~ n)

g7 = ggplot(fit_identify_result_type, aes(x = method, y = F1_score, color = type)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "F1_score",x = "method", y = "F1_score") +
  facet_grid(ratio ~ n)

g8 = ggplot(fit_identify_result_type, aes(x = method, y = accuracy, color = type)) + 
  geom_boxplot() +
  theme(legend.position = "right") +
  labs(title = "Accuracy",x = "method", y = "accuracy") +
  facet_grid(ratio ~ n)


# Beta estimation of RMSE

# RMSE of beta for n = 100
g9 = ggplot(fit_beta_result_ratio %>% filter(n == 100), aes(x = method, y = RMSE)) + 
  geom_boxplot(color = "red", fill = "orange", alpha = 0.5) +
  theme(legend.position = "right") +
  labs(title = "beta RMSE when n = 100",x = "method", y = "RMSE") +
  coord_cartesian(ylim = c(0, 15)) +
  facet_grid(~ ratio)

# RMSE of beta for n = 500
g10 = ggplot(fit_beta_result_ratio %>% filter(n == 500), aes(x = method, y = RMSE)) + 
  geom_boxplot(color = "red", fill = "orange", alpha = 0.5) +
  theme(legend.position = "right") +
  labs(title = "beta RMSE when n = 500",x = "method", y = "RMSE") +
  coord_cartesian(ylim = c(0, 0.6)) +
  facet_grid(~ ratio)

# RMSE of beta for n = 2000
g11 = ggplot(fit_beta_result_ratio %>% filter(n == 2000), aes(x = method, y = RMSE)) + 
  geom_boxplot(color = "red", fill = "orange", alpha = 0.5) +
  theme(legend.position = "right") +
  labs(title = "beta RMSE when n = 2000",x = "method", y = "RMSE") +
  coord_cartesian(ylim = c(0, 0.4)) +
  facet_grid(~ ratio)


# Beta estimation of Variance

# Variance of beta for n = 100
g12 = ggplot(fit_beta_result_ratio %>% filter(n == 100), aes(x = method, y = Variance)) + 
  geom_boxplot(color = "red", fill = "orange", alpha = 0.5) +
  theme(legend.position = "right") +
  labs(title = "beta Variance when n = 100",x = "method", y = "Variance") +
  coord_cartesian(ylim = c(0, 150)) +
  facet_grid(~ ratio)

# Variance of beta for n = 500
g13 = ggplot(fit_beta_result_ratio %>% filter(n == 500), aes(x = method, y = Variance)) + 
  geom_boxplot(color = "red", fill = "orange", alpha = 0.5) +
  theme(legend.position = "right") +
  labs(title = "beta Variance when n = 500",x = "method", y = "Variance") +
  coord_cartesian(ylim = c(0, 50)) +
  facet_grid(~ ratio)

# Variance of beta for n = 2000
g14 = ggplot(fit_beta_result_ratio %>% filter(n == 2000), aes(x = method, y = Variance)) + 
  geom_boxplot(color = "red", fill = "orange", alpha = 0.5) +
  theme(legend.position = "right") +
  labs(title = "beta Variance when n = 2000",x = "method", y = "Variance") +
  coord_cartesian(ylim = c(0, 150)) +
  facet_grid(~ ratio)
```

```{r, echo = FALSE, fig.cap = "Model Complexity", fig.height = 8,fig.width = 7}
g_1
```

```{r, echo = FALSE, fig.cap = "Overall Classification Performance", fig.height = 8,fig.width = 9}
(g1 + g2) / (g3 + g4)
```

```{r, echo = FALSE, fig.cap = "Classification Performance by Signals", fig.height = 8,fig.width = 9}
(g5 + g6) / (g7 + g8)
```

```{r, echo = FALSE, fig.cap = "Parameter Estimation Performance", fig.height = 8,fig.width = 9}
(g9 + g12) / (g10 + g13) / (g11 + g14)
```
