---
title: "Project 1-Topic 3 Report"
author: "Hongjie Liu, Jiajun Tao, Shaohan Chen"
date: "02/27/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gridExtra)
library(tidyverse)
```

## Objectives

When dealing with high-dimensional data

## Statistical Methods Studied

### Stepwise Forward Selection

Step-wise forward selection 

### LASSO

LASSO regression 

## Scenarios Investigated

Several parameters 


## Methods for Generating Data

Data is generated using a 4-step method:

## Performance Measures

### Ability of Model to Identify Weak and Strong Predictors

One method of


### Effect of Missing Weak Predictors

On the other hand
## Results

According to Table 2, forward stepwise always includes more predictors than the number of true predictors in original data while LASSO is more selective.

### Scenarios 1-4: Varying total number of weak predictors 

*Type I error and power*

While both methods do are decent at selecting true predictors for the model, LASSO is much better at excluding null predictors than forward selection (Table 3). 


*Coefficient estimation*

Although the median SE of the coefficient estimates for forward stepwise is lower than LASSO, its MSE is higher when there are few true weak predictors in the data.
### Scenarios 4-8: Varying degree of correlation

*Type I error and power*

As the correlation of the WBC predictors increases, the type I error for forward stepwise and LASSO remain fairly constant around 0.35 and 0.05, respectively. Meanwhile, the power of forward selection increases, while the power of LASSO decreases slightly. (Figure 6)  

*Coefficient estimation*

Although forward stepwise selection includes more non-null predictors on average (Table 3), it has increasing and higher MSE compared with LASSO, showing a growing trend when correlation becomes larger. The MSE of LASSO is controlled under 0.1, whereas forward selection has an inflated MSE (over 0.2) when the correlation gets very high. 

## Conclusion

### General

Forward stepwise selection always includes more predictors than LASSO, and the model size given by each method increases as the number of total true predictors in the original data increases.

### Varying total number of weak predictors

The SE variance for forward stepwise is much larger than LASSO when the number of weak predictors is small

### Varying degree of correlation

With increased correlation

## Discussion

As the degree of correlation increases


\pagebreak

## Figures and Tables

```{r figs, echo=FALSE, warning=FALSE, message=FALSE}
n=1
```